---
title: "Project 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/oli/Uni-Stuff/2022/Sem 2/Data Mining/Projects/Project 2/Data/initial")
```

```{r, include=FALSE}
setwd("/Users/oli/Uni-Stuff/2022/Sem 2/Data Mining/Projects/Project 2/Data/initial")
library(tidyverse)
library(caret)
library(randomForest)
library(ggplot2)
library(scales)
library(Hmisc)
library(plyr)
library(dplyr)

getwd()
```

Load in the data
```{r}
getwd()
transaction <- read.csv("train_transaction.csv", header=TRUE) %>% na_if("")
identity <- read.csv("train_identity.csv", header=TRUE) %>% na_if("")
attach(transaction)
attach(identity)
```

Merge identity and transaction data
```{r}
data <- merge(transaction,identity,by="TransactionID", all.x=TRUE)
data$isFraud <- as.factor(data$isFraud)
```

Drop columns with more than 80% NA vlues
```{r}
na = colSums(is.na(data))/nrow(data)
criteria = 0.8
data = data[,!sapply(data, function(x) mean(is.na(x)))>criteria]

```

Identify two anomalies in Transaction amount and look at box plots of TransactionAmt for fraudulent and non-fraudulent transactions. 
```{r}
summary(data$TransactionAmt)
boxplot(data$TransactionAmt)
data <- data[data$TransactionAmt<10000,]

y<-data[,2]
x<-data[,4]
featurePlot(x,y,"box")
```

Only 3% of data is fraudulent
```{r}
100*table(data$isFraud)/length(data$isFraud)

```

Feature engineering of TransactionDT column to get it into hour of the day. Clearly the fraudulent transactions are taking place at similar points in the day. 
```{r}
data['hour']=(floor(data['TransactionDT']/3600))%%24
fraction<-aggregate(as.numeric(data$isFraud)-1, list(data$hour), FUN=mean)
plot(x=fraction$Group.1, y=fraction$x, xlab="Hour",ylab="Fraction of all transactions that were fraudulent",type="l")
ggplot(data, aes(x=hour, fill=isFraud))+geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge") + scale_y_continuous(labels=percent)

```


Reduce the number of categories for P and R email domain. 
```{r}

data <- data %>% mutate(P_emaildomain = fct_collapse(P_emaildomain, "google"=c("gmail","gmail.com"), "yahoo"=c("yahoo.co.jp","yahoo.co.uk","yahoo.com","yahoo.com.mx","yahoo.de","yahoo.es","yahoo.fr"),"Microsoft"=c("hotmail.co.uk","hotmail.com","hotmail.de","hotmail.es","hotmail.fr","live.com","live.com.mx","live.fr","outlook.com","outlook.es"),other_level="other"))


ggplot(data, aes(x=P_emaildomain, fill=isFraud))+geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge") + scale_y_continuous(labels=percent)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

data <- data %>% mutate(R_emaildomain = fct_collapse(R_emaildomain, "google"=c("gmail","gmail.com"), "yahoo"=c("yahoo.co.jp","yahoo.co.uk","yahoo.com","yahoo.com.mx","yahoo.de","yahoo.es","yahoo.fr"),"Microsoft"=c("hotmail.co.uk","hotmail.com","hotmail.de","hotmail.es","hotmail.fr","live.com","live.com.mx","live.fr","outlook.com","outlook.es"),other_level="other"))
```

Remove all V columns
```{r}
data <- data[,!grepl("V",names(data))]
```

Reduce number of categories in id_31
```{r}
n_distinct(data$id_31)

data <- data %>% mutate(id_31 = fct_collapse(id_31, "chrome"=c("chrome","chrome 63.0 for ios","chrome 43.0 for android","chrome 46.0 for android","chrome 49.0","chrome 49.0 for android","chrome 50.0 for android","chrome 51.0","chrome 51.0 for android","chrome 51.0 for android","chrome 52.0 for android","chrome 53.0 for android","chrome 54.0 for android","chrome 55.0","chrome 55.0 for android","chrome 56.0","chrome 56.0 for android","chrome 57.0","chrome 57.0 for android","chrome 58.0","chrome 58.0 for android","chrome 59.0","chrome 59.0 for android","chrome 60.0","chrome 60.0 for android","chrome 61.0","chrome 61.0 for android","chrome 62.0","chrome 62.0 for android","chrome 62.0 for ios","chrome 63.0","chrome 63.0 for android","chrome 64.0","chrome 64.0 for android","chrome 64.0 for ios","chrome 65.0","chrome 65.0 for android","chrome 65.0 for ios","chrome 66.0","chrome 66.0 for android","chrome 66.0 for ios","chrome 67.0","chrome 67.0 for android","chrome 69.0","chrome generic","chrome generic for android"),"opera"=c("opera","opera 49.0","opera 51.0","opera 52.0","opera 53.0","opera generic"),"samsung"=c("samsung","samsung browser 3.3","samsung browser 4.0","samsung browser 4.2","samsung browser 5.2","samsung browser 5.4","samsung browser 6.2","samsung browser 6.4","samsung browser 7.0","samsung browser generic","Samsung/SCH","Samsung/SM-G531H","Samsung/SM-G532M"),"firefox"=c("firefox","firefox 47.0","firefox 48.0","firefox 52.0","firefox 55.0","firefox 56.0","firefox 57.0","firefox 58.0","firefox 59.0","firefox 60.0","firefox generic","firefox mobile 61.0"),"edge"=c("edge","edge 13.0","edge 14.0","edge 15.0","edge 16.0","edge 17.0"),"safari"=c("safari","safari 10.0","safari 11.0","safari 9.0","safari generic","mobile safari","mobile safari 10.0","mobile safari 11.0","mobile safari 8.0","mobile safari 9.0", "mobile safari generic","mobile safari uiwebview"),"NA"="NA",other_level = "Other"))

ggplot(data, aes(x=id_31, fill=isFraud))+geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge") + scale_y_continuous(labels=percent)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Based on correlation with isFraud only keep certain id columns. Drop columns with too many categories in them. 
```{r}
drop <- c("TransactionDT", "DeviceInfo","id_28","card1","card2","card5","addr1","id_01","id_02","id_05","id_06","id_11","id_13","id_17","id_19","id_20","id_35","id_36","id_37","id_38")
data=data[,!names(data) %in% drop]
```

For remaining columns, ensure they are of correct data type and reduce number of categories. 
```{r}
data[sapply(data,is.character)]<-lapply(data[sapply(data,is.character)],as.factor)


data$card3 <- as.factor(data$card3)
data <- data %>% mutate(card3 = fct_collapse(card3, "150"="150","185"="185",other_level="other"))

data$addr2 <- as.factor(data$addr2)
data <- data %>% mutate(addr2 = fct_collapse(addr2, "87"="87",other_level = "other"))

```

Transform logical into factors.
```{r}
data$M1 <- as.factor(as.numeric(data$M1))
data$M2 <- as.factor(as.numeric(data$M2))
data$M3 <- as.factor(as.numeric(data$M3))
data$M5 <- as.factor(as.numeric(data$M5))
data$M6 <- as.factor(as.numeric(data$M6))
data$M7 <- as.factor(as.numeric(data$M7))
data$M8 <- as.factor(as.numeric(data$M8))
data$M9 <- as.factor(as.numeric(data$M9))
```

Imputation of mean and mode values to replace all NAs. 
```{r}
data[sapply(data, is.numeric)] <- lapply(data[sapply(data, is.numeric)], function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))

data$card3<-impute(data$card3,mode)
data$card4<-impute(data$card4,mode)
data$card6<-impute(data$card6,mode)
data$addr2<-impute(data$addr2,mode)
data$P_emaildomain<-impute(data$P_emaildomain,mode)
data$R_emaildomain<-impute(data$R_emaildomain,mode)
data$M4<-impute(data$M4,mode)
data$M1<-impute(data$M1,mode)
data$M2<-impute(data$M2,mode)
data$M3<-impute(data$M3,mode)
data$M5<-impute(data$M5,mode)
data$M6<-impute(data$M6,mode)
data$M7<-impute(data$M7,mode)
data$M8<-impute(data$M8,mode)
data$M9<-impute(data$M9,mode)
data$id_12<-impute(data$id_12,mode)
data$id_15<-impute(data$id_15,mode)
data$id_16<-impute(data$id_16,mode)
data$id_29<-impute(data$id_29,mode)
data$id_31<-impute(data$id_31,mode)
data$DeviceType<-impute(data$DeviceType,mode)

write.csv(data,"Ailsadatafinal.csv")

```

The same analysis is applied to the test data

```{r}
train <- read.csv("Ailsadatafinal.csv",header=TRUE)
testing1 <- read.csv("Ailsatestfinal.csv",header=TRUE)
```

Training data adjusted to make sure the columns are of the correct class
```{r}
train$isFraud <- as.factor(train$isFraud)
train[sapply(train,is.character)]<-lapply(train[sapply(train,is.character)],as.factor)
train$M1 <- as.factor(train$M1)
train$M2 <- as.factor(train$M2)
train$M3 <- as.factor(train$M3)
train$M5 <- as.factor(train$M5)
train$M6 <- as.factor(train$M6)
train$M7 <- as.factor(train$M7)
train$M8 <- as.factor(train$M8)
train$M9 <- as.factor(train$M9)

drop <- c("X","TransactionID")
train=train[,!names(train) %in% drop]
```

Random forest fit (mtry=p/3). There are 48 variables that the Random Forest will fit to. 
```{r}
set.seed(0597)
index <- sample(nrow(train), size = floor(0.5*nrow(train)))


RF2 <- randomForest(isFraud ~. , data=train, subset=index, mtry=16, ntree=100, importance=TRUE)

RF2

varImpPlot(RF2)
plot(RF2)
```

Ensure the testing set is in correct format
```{r}
X <- c("TransactionID","D6","D13","D14")
testing<-testing1[,!names(testing1) %in% X]

testing[sapply(testing,is.character)]<-lapply(testing[sapply(testing,is.character)],as.factor)
testing$M1 <- as.factor(testing$M1)
testing$M2 <- as.factor(testing$M2)
testing$M3 <- as.factor(testing$M3)
testing$M5 <- as.factor(testing$M5)
testing$M6 <- as.factor(testing$M6)
testing$M7 <- as.factor(testing$M7)
testing$M8 <- as.factor(testing$M8)
testing$M9 <- as.factor(testing$M9)

levels(testing$card6) <- levels(train$card6)

```

Make predictions for test data. Submit vector is what was entered into Kaggle. 
```{r}
testing1$isFraud <- predict(RF2,newdata=testing,type="prob")[,2]


y <- c("TransactionID","isFraud")
submit <- testing1[y]

```

